Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-06 paper/paper.qmd
fix typo and update decimal

### States included in anaylsis

@@

```{r}
#| label: fig-avg-pct-state
#| fig-cap: Historical State Voting Trends Are Maintined in 2024
#| echo: false
#| warning: false
#| 
cleaned_data <- cleaned_data %>% mutate(candidate_name =
  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))

@@

```{r}
#| label: fig-polldate-canidate-pct
#| fig-cap: Both Canidate Percentages Are Gaining Support Closer to the Election
#| echo: false
#| warning: false

@@

Next, we set up a bayesian model as the following in accordance to @sec-modelsetup:

@@

```{r}
#| label: fig-bayesian-model-actual-predict
#| fig-cap: "The MLR Model Accruatley Predicts the Testing Dataset"
#| echo: false
#| warning: false
plot(test_data$pct, colMeans(predictions), 
     xlab = "Actual Voter Percentage Values", 
     ylab = "Predicted Voter Percentage Values", 
     main = "Multiple Predictors Kamala Harris Bayesian Model")
abline(0, 1, col = "red")
```

@@

## Harris Voter Prediction Model 
```{r}
#| label: fig-bayesian-model-CI
#| fig-cap: "Kamala Harris is Predicted to get Approximatley 52% of Popular Vote"
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95, 
     xlab = "Kamala Harris Popular Vote Percentage",
     title = "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model")
```

@@

predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))

@@

  avg_predicted_pct_state_harris <- mean(predictions_state_harris)
  avg_predicted_pct_state_trump <- mean(predictions_state_trump)

@@

# Overview of American Electon System

@@

```{r}
#| label: fig-trumpmodelci
#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
#| echo: false
#| warning: false
plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
```

@@

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "Bayesian MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@@

kable(head(result_summary, 10), col.names = c("Acutal Harris Percentage", 
                    "Predicted Percentage", 
                    "Lower Bound Credible Interval",
                    "Upper bound Credible Interval"))

>>>>

### States included in Analysis

@@

```{r}
#| label: fig-avg-pct-state
#| fig-cap: Historical State Voting Trends Are Maintained in 2024
#| echo: false
#| warning: false
#| 
cleaned_data <- cleaned_data %>% mutate(candidate_name =
  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))

@@

```{r}
#| label: fig-polldate-canidate-pct
#| fig-cap: Both Candidate Percentages Are Gaining Support Closer to the Election
#| echo: false
#| warning: false

@@

Next, we set up a Bayesian model as the following in accordance to @sec-modelsetup:

@@

```{r}
#| label: fig-bayesian-model-actual-predict
#| fig-cap: "The MLR Model Accurately Predicts the Testing Dataset"
#| echo: false
#| warning: false
plot(test_data$pct, colMeans(predictions), 
     xlab = "Actual Voter Percentage Values", 
     ylab = "Predicted Voter Percentage Values", 
     main = "Multiple Predictors Kamala Harris Bayesian Model")
abline(0, 1, col = "red")
```

@@

## Harris Voter Prediction Model 
```{r}
#| label: fig-bayesian-model-CI
#| fig-cap: "Kamala Harris is Predicted to get Approximately 52% of the Popular Vote"
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95, 
     xlab = "Kamala Harris Popular Vote Percentage",
     title = "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model")
```

@@

predicted_means = round(colMeans(predictions2), 2)
predicted_intervals = round(apply(predictions2, 2, quantile, probs = c(0.025, 0.975)), 2)

@@

  avg_predicted_pct_state_harris <- round(mean(predictions_state_harris), 2)
  avg_predicted_pct_state_trump <- round(mean(predictions_state_trump), 2)

@@

# Overview of American Election System

@@

```{r}
#| label: fig-trumpmodelci
#| fig-cap: "Donald Trump Expected to Receive Approximately 45% of the Vote"
#| echo: false
#| warning: false
plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
```

@@

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "Bayesian MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = round(colMeans(predictions2), 2)
predicted_intervals = round(apply(predictions2, 2, quantile, probs = c(0.025, 0.975)), 2)
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@@

kable(head(result_summary, 10), 
      col.names = c("Actual Trump Percentage", 
                    "Predicted Percentage", 
                    "Lower Bound Credible Interval",
                    "Upper Bound Credible Interval"))
```
